---
title: "MacroRegions"
---

```{r, echo=FALSE,message=FALSE,error = FALSE}
library(knitr)
library(dplyr, quietly=TRUE)
library(data.table)
library(FactoMineR)

knitr::opts_chunk$set(cache = FALSE,
                      echo = FALSE,
                      comment = "")
```



## Data

We laod an **hypercube** where the text of news has been removed and where we keep only the number of *tags* or proportion of *news* speaking from one or several regions (*where1*, *where2*), by media (*who*) and by time period (*when*)

```{r}
# Load Hypercube and select period
hc_reg<-readRDS("hypercube/hc_reg_reg_day.RDS") %>% filter(where1 !="_no_")
hc_reg %>% filter(is.na(when)==F, when < as.Date("2021-07-01"))


# Define weights correction for equal contribution of media
wgt<-hc_reg[,list(coeff = sum(tags)), list(who)]  %>% 
      mutate(coeff=mean(coeff)/coeff)
hc_reg<-left_join(hc_reg,wgt)
hc_reg$news_wgt<-hc_reg$news*hc_reg$coeff

# Load table of label and choose language
reg_def<-read.table("dict/worldgeo_def_V2.csv", sep=";",quote = '"', encoding = "UTF-8",header=T)
tab_def<-reg_def %>% filter(lang=="en") %>% select(code,type,label)
tab_def<-tab_def[duplicated(tab_def$label)==F,]
```

## Top 50 regions in full corpus


### Unweighted

We can propose firstly a table of top entities in the whole corpus of newspapers with index 100 for the first entity.

```{r}
# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(where1)] 
df<-merge(tab_def,df,by.x="code",by.y="where1",all.x=F,all.y=T)
df<-df[order(df$nb, decreasing = T),]
row.names(df)<-1:dim(df)[1]
df$index<-100*df$nb/max(df$nb)

kable(head(df,50), digits=c(NA,NA,NA,0,2),row.names = T)
```

### Weighted


```{r}
df<-hc_reg[where1 !="_no_",list(nb = sum(news_wgt)), list(where1)] 
df<-merge(tab_def,df,by.x="code",by.y="where1",all.x=F,all.y=T)
df<-df[order(df$nb, decreasing = T),]
row.names(df)<-1:dim(df)[1]
df$index<-100*df$nb/max(df$nb)

kable(head(df,50), digits=c(NA,NA,NA,0,2),row.names = T)
```


## Top 10 regions by media

### German newspapers - Top 10 regions

```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
     filter(where1 != "Q828")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6)=="DEU", rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")

res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:10,2]
names(tab1) <- c("Rank")
tab2<-res[1:10,c(3,4)]
names(tab2)<-c(paste("Frankf. Allg. Zeit.","Region",sep="_"),paste("FAZ","pct",sep=" "))
tab3<-res[11:20,c(3,4)]
names(tab3)<-c(paste("Süd. Zeit.","Region",sep="_"),paste("Süd. Zeit.","pct",sep=" "))
tab<-cbind(tab1,tab2,tab3)

kable(tab,digits=1, row.names = F, 
      col.names = c("Rank","Frankf. Allg. Zeit.","index","Südeutsche Zeit.","index"))

```

### French newspapers - Top 10 regions

```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
     filter(where1 != "Q828")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6)=="FRA", rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")

res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:10,2]
names(tab1) <- c("Rank")
tab2<-res[1:10,c(3,4)]
names(tab2)<-c(paste("Figaro","Region",sep="_"),paste("Figaro","pct",sep=" "))
tab3<-res[11:20,c(3,4)]
names(tab3)<-c(paste("Le Monde","Region",sep="_"),paste("Le Monde","pct",sep=" "))
tab<-cbind(tab1,tab2,tab3)

kable(tab,digits=1, row.names = F,
      col.names = c("Rank","Le Figaro","index","Le Monde","index"))

```

### UK newspapers - Top 10 regions

```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
     filter(where1 != "Q828")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6)=="GBR", rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")

res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:10,2]
names(tab1) <- c("Rank")
tab2<-res[1:10,c(3,4)]
names(tab2)<-c(paste("Guardian","Region",sep="_"),paste("Guardian","pct",sep=" "))
tab3<-res[11:20,c(3,4)]
names(tab3)<-c(paste("Daily Telegraph","Region",sep="_"),paste("Daily Telegraph","pct",sep=" "))
tab<-cbind(tab1,tab2,tab3)

kable(tab,digits=1, row.names = F, 
      col.names = c("Rank","The Guardian","index","The Daily Telegraph","index"))

```

### Irish newspapers - Top 10 regions

```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
     filter(where1 != "Q828")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6) %in% c("IRL","NIR"), rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")

res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:10,2]
names(tab1) <- c("Rank")
tab2<-res[1:10,c(3,4)]
names(tab2)<-c(paste("Irish Times","Region",sep="_"),paste("Irish Times","pct",sep=" "))
tab3<-res[11:20,c(3,4)]
names(tab3)<-c(paste("Belfast Telegraph","Region",sep="_"),paste("Belfast Telegraph","pct",sep=" "))
tab<-cbind(tab1,tab2,tab3)

kable(tab,digits=1, row.names = F, 
      col.names = c("Rank","Irish Times","index","Belfast Telegraph","index"),)

```





### Turkish newspapers - Top 10 regions


```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
#     filter(where1 != "Q828")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6)=="TUR", rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")


res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:10,2]
names(tab1) <- c("Rank")
tab2<-res[1:10,c(3,4)]
names(tab2)<-c(paste("Cumhuryet","Region",sep="_"),paste("Cumhuryet","pct",sep=" "))
tab3<-res[11:20,c(3,4)]
names(tab3)<-c(paste("Yeni Savak","Region",sep="_"),paste("Yeni Savak","pct",sep=" "))
tab<-cbind(tab1,tab2,tab3)

kable(tab,digits=1, row.names = F, 
      col.names = c("Rank","Cumhurryet","index","Yeni Savak","index"))

```

### Algerian newspapers

```{r}

# Compute
df<-hc_reg[where1 !="_no_",list(nb = sum(news)), list(who, where1)] %>% 
     group_by(who) %>%
     filter(where1 != "LA_sahara")%>%
     mutate(pct = 100*nb/max(nb),
            rnk = rank(-nb))

df_sel <- df %>% filter(substr(who,4,6)=="DZA", rnk < 11)
df_sel<-merge(df_sel,tab_def,by.x="where1",by.y="code")

res <- df_sel %>% filter(rnk < 11) %>% select(who, rnk,label, pct) %>% mutate(who=substr(who,4,12))
res<-res[order(res$who, res$rnk),]
tab1<-res[1:11,2]
names(tab1) <- c("Rank")
tab2<-res[1:11,c(3,4)]
names(tab2)<-c("Al Nahar ","pct1")
tab3<-res[12:22,c(3,4)]
names(tab3)<-c("El Kahbar","pct2")


tab<-cbind(tab1,tab2,tab3)


kable(tab,digits=c(0,1,1,1,1,1), row.names = F,
      col.names = c("Rank","Al Nahar","index","El Khabar","index"))

```




## Synthesis by correspondance analysis and Hierarchical Clustering

The synthesis is realized with regions

### Factor 1-2


```{r}
# Matrix
reg_med <-hc_reg[where1 !="_no_",list(nb = sum(news_wgt)), list(where1, who)] %>%
  dcast(formula = where1~who, value.var = "nb",fill = 0) 
reg_med<-merge(reg_med,tab_def, by.x="where1",by.y="code",all.x=T,all.y=F)
reg_med$label[reg_med$label=="Machrek"]<-"Maghreb"
mat<-as.matrix(reg_med[,2:13])
row.names(mat)<-reg_med$label

# Filter ambiguous units
#mat<-mat[row.names(mat) != "Sahara",]
##mat<-mat[row.names(mat) != "Asia Minor",]
#mat<-mat[row.names(mat) != "Southern Africa",]
#mat<-mat[row.names(mat) != "Europe",]
#mat<-mat[row.names(mat) != "European Union",]


# Select units 
sel<-mat[apply(mat,1,sum)>10,]

# Exclude units mentionned by less than 3 media
sel <- sel[apply(sel>3,1,sum)>2,]


afc <- CA(sel, graph = F)
#library(explor)
#explor(afc)
res <- explor::prepare_results(afc)
explor::CA_var_plot(res, xax = 1, yax = 2, lev_sup = FALSE, var_sup = FALSE,
    var_sup_choice = , var_hide = "None", var_lab_min_contrib = 4, col_var = "Position",
    symbol_var = NULL, size_var = "Contrib", size_range = c(52.5, 700), labels_size = 10,
    point_size = 56, transitions = FALSE, labels_positions = "auto", xlim = c(-2,
        1), ylim = c(-1, 2))

```

### Factors 3-4

```{r}
res <- explor::prepare_results(afc)
explor::CA_var_plot(res, xax = 3, yax = 4, lev_sup = FALSE, var_sup = FALSE,
    var_sup_choice = , var_hide = "None", var_lab_min_contrib = 4, col_var = "Position",
    symbol_var = NULL, size_var = "Contrib", size_range = c(52.5, 700), labels_size = 10,
    point_size = 56, transitions = FALSE, labels_positions = "auto", xlim = c(-0.924,
        2.01), ylim = c(-1.5, 1.43))
```

### Cluster analysis (world regions)

```{r}
cah1 <- HCPC(afc,nb.clust = 5,graph = FALSE)
plot.HCPC(cah1,choice="tree")
```

### Cluster analysis (medias)

```{r}
cah2 <- HCPC(afc,nb.clust = 4,graph = FALSE,cluster.CA = "columns")
plot.HCPC(cah2,choice="tree")
```

## Synthesis by chi-square & heatmap

An alternative approach based on the computation of chi-square and the classification of residuals (censoredd between -3 and +3)

```{r,fig.height=8, fig.width =6}
library(pheatmap)
# Select units 
sel<-mat[apply(mat,1,sum)>24,]

# Exclude units mentionned by less than 5 media
sel <- sel[apply(sel>3,1,sum)>3,]

x<-chisq.test(sel)
res<-as.matrix(x$residuals)
exp<-as.matrix(x$expected)
res[exp<2]<-0
res[res>3]<-3
res[res< -3]<- -3

pheatmap(res, cutree_rows = 6,cutree_cols = 6)


```




 







